---
title: "Iris Classfication Application"
author: "LI CHENYANG"
logo: iris.png
date: "December 24, 2017"
output: ioslides_presentation
widscreen: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(caret)
library(ggplot2)
library(plotly)
```

## Iris Classification

* This app is created to classify iris, based on the attributes specified in "Edgar Anderson's Iris Data":

    **Sepal Length,   Sepal Width,  Petal Length,  Petal Width**

* When user keys in the values for the attributes, a prediction will be made. 
* For demonstration purpose, we will use the values below as example. 
    + Sepal Length = 6cm
    + Sepal Width = 3cm
    + Petal Length = 6cm
    + Petal Width = 2cm 


## Apps Demonstration: Prediction
### Creating the model
```{r, echo = TRUE, warning=FALSE, message=FALSE}
irisModel <- train(Species~ .,data=iris,method="rf",prox=TRUE)
```
### Loading input into the model
```{r, echo = TRUE, warning=FALSE, message=FALSE}
inputData <- data.frame(Sepal.Length = 6, Sepal.Width = 3,
                        Petal.Length = 6, Petal.Width = 2)
result <- predict(irisModel, inputData)
result
```
## Apps Demonstration: 1st Scatter Plots
###### The measurement of new flower is shown by the diamond shape.
```{r, echo = FALSE, warning=FALSE, message=FALSE}
p1 <- ggplot(data = iris, aes(x = Petal.Length, 
                              y = Sepal.Length, color = Species
                              )) + geom_point()
inputData$Species <- result
p1 <- p1 + geom_point(data = inputData, shape = 23, size = 4)
ggplotly(p1)
```

## Apps Demonstration: 2nd Scatter Plots
###### This figure plots the "width" variables against each other.
```{r, echo = FALSE, warning=FALSE, message=FALSE}
p2<- ggplot(data = iris, aes(x = Petal.Width, 
                              y = Sepal.Width, color = Species
                              )) + geom_point()
inputData$Species <- result
p2 <- p2 + geom_point(data = inputData, shape = 23, size = 4)
ggplotly(p2)
```
