---
title: "Stats_Inference_Project_PartII"
author: "LI CHENYANG"
date: "December 25, 2015"
output: pdf_document
---

## Overview
- In the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package. 

## 1. Basic Exploratory data analyses

- In this case, I plot six histograms, with each representing one possible combination between supp and dose factors. 

- From the graph below, we can see roughly see that
    
    
    1.OJ may be a better supplement than VC when it comes to teeth growth. 
    
    2.The higher the dose, the longer the teeth will grow. However I am more interested in the extent of growth when the dosage increases. By the law of marginal diminishing return, I would guess that increasing dosage from 1 to 2 will not necessarily yeild a better growth than increasing dosage from 0.5 to 1. 

```{r, echo=FALSE, message = FALSE}
library(ggplot2)
library(plyr)
library(dplyr)
RawData <- ToothGrowth
RawData$dose <- factor(RawData$dose)
g <- ggplot(data = RawData, aes(x=len, fill = dose)) + geom_histogram(binwidth=0.8)
g + facet_grid(supp~dose)
```


## 2. Provide a basic summary of the data.

- The data frame has 60 observation made by the same 10 guinea pigs. 
- The data fram has 3 columns, which are len, supp and dose
- len represent the length of the teeth
- supp represent the supplement type
- dose represent dosage amount for the supplement

perhaps a statistic summary will help us understand the big picture

```{r echo=FALSE}
RawData <- group_by(RawData, supp, dose)
summarize(.data = RawData, mean_len = mean(len), var_len = var(len))

```


## 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

### Compare tooth growth by dose 

- I would like to compare the growth of teeth between two scenarios. First scenarios is when dosage is raised from 0.5 to 1, and second scenarios is when dosage raised from 1 to 2

- **Null Hypotheses**
- H0: mean of difference in scenario 1 = mean of differen in scenario 2

- set alpha as 0.05 

- I consider dose = 0.5 as the base level. 
 
- **Assumption**: 
- The growth in tooth is not affected by factors not mentioned in the datasets
- The guinea pig has not done anything to shorten its teeth during the process. 
- All the observations made are based on the original 10 guinea pigs. 



- We only use data from the VC group

```{r}
g10 <- RawData$len[11:20] - RawData$len[1:10]
g21 <- RawData$len[21:30] - RawData$len[11:20]
t.test(g10, g21, paired = TRUE)$p.value

```

- The p-value = 0.829 > 0.05
- Therefore I do not reject the null hypothesis
- This result suggest that increasing the dose from 1 to 2 renders similar growth result as when increasing dose from 0.5 to 1. 


### Compare tooth growth by changing supp

- compare growth when changing supp for dose group 0.5, calculate confidence interval 
```{r}
g0_s <- RawData$len[1:10]- RawData$len[31:40]
g1_s <- RawData$len[11:20]- RawData$len[41:50]
g2_s <- RawData$len[21:30]- RawData$len[51:60]
t.test(g0_s)$conf

```

- The confidence interval (95%, 2 side) is [-9.24, -1.26], entirely below zero. 
- This suggest that changing VC to OJ for dose 0.5 totally increased the tooth length
  
```{r}

t.test(g1_s)$conf
```

- The confidence interval (95%, 2 side) is [-9.908089 -1.951911], entirely below zero. 
- This suggest that changing VC to OJ for dose 1 totally increased the tooth length
  
 
```{r}

t.test(g2_s)$conf
```

- The confidence interval (95%, 2 side) is [-4.168976  4.328976], evenly distributed around 0.
- In this case the change of supp *may not have made a difference*.



## 4. State your conclusions and the assumptions needed for your conclusions. 

- **Conclusion 1**, increasing dosage from 0.5 to 1 generates similar tooth growth as increasing dosage from 1 to 2. 

- ** Conlusion 2**, Changing supplement from VC to OJ helps tooth growth when the dosage is low. When dosage reaches 2, the change of supplement does not affect the tooth growth very much. 




